The Blue Obelisk five years on

### 2011-03-08 paper.tex
Added some text on toolkit progress

...something about toolkits...

Although feature-rich and robust cheminformatics toolkits are useful in and of themselves, they can also be seen as providing a base layer on which tools and applications can be built. Now that these toolkits have reached an acceptable level of stability and functionality, an increasing amount of software is being developed using these toolkits. As an example of the interaction between the commercial and open source worlds, Silicos (a Belgian company that provides services in the area of cheminformatics XXXX rephrase), has released several command line applications based on Open Babel; Pharao for pharmacophore generation (Sept 2010?), Sieve for filtering by molecular property, .....

It is worth noting that Pharao is the first Open Source pharmacophore program (\emph{Is this a reasonable statement?}) (although pharmacophore-related functionality has been available in the CDK and RDKit for some time).

>>>>

Open Source toolkits for cheminformatics have now existed for nearly
ten years. During this period, some toolkits were developed form
scratch, whereas others were made Open Source by releasing in-house
codebases under liberal licenses. The primary toolkits currently in
active development are the Chemistry Development Kit, OpenBabel and
RDKit. These projects support a variety of languages natively (Java,
C++, Python), but by virtue of language wrappers, they can easily be
used from languages that they were not originally developed in. For
example, the CDK can be accessed from Python and OpenBabel can be
accesed from Java via SWIG wrappers. One important aspect of the
toolkits noted here is that they originated for the support of desktop
and server-side applications.

More recently, a number of workers have developed light-weight
toolkits that are targeted towards web applications. One example of
this form is the Chemdoodle library from XXX. The focus of these
toolkits is to support client-side cheminformatics functionality,
allowing for rich, chemically-aware browser-based applications.

In terms of functionality, most Open Source libraries support commonly
ued cheminformatics features. At the same time, individual toolkits
have their own strenghts. For example, the CDK has an extensive set of
molecular descriptors useful for QSAR modeling applictions, but lacks
support for 3D structure generation. In contrast OpenBabel, supports
3D structure generation and optimization using a variety of
forcefields, but lacks a built in set of molecular descriptors. RDKit
on the other hand supports both features and in general is the most
comprehensive Open Source toolkit in terms of functionality. Table XXX
summarizes the features of these three toolkits.

Although feature-rich and robust cheminformatics toolkits are useful
in and of themselves, they can also be seen as providing a base layer
on which tools and applications can be built. Now that these toolkits
have reached an acceptable level of stability and functionality, an
increasing amount of software is being developed using these
toolkits. A number of Open Source applications including Bioclipse,
XXX have been developed.

More interestingly, commercial groups have released Open Source tools
based on these toolkits. An recent example of such an interaction 
is Silicos (a Belgian company that provides services
in the area of cheminformatics XXXX rephrase), has released several
command line applications based on Open Babel. For example, the Pharao
tool released by them is a comprehensive solution for pharmacophore
searching that provides extensive support for a variety of
pharmacophore searches. Other tools released by them include Seive for
fitering by molecular property and XXX. Another example of Open Source
tools originating from commercial groups is the ChemCraft tool from
Molecular Networks GmBH, that does XXX.
